Inequality in socioemotional skills: a cross-cohort comparison

### 2019-08-25 codes/R/paper/text/paper_jpube_2datamethods.tex
text updates by Gabriella

Another complication arises from the fact that, differently from the British Cohort Study, the Millennium Cohort Study has a stratified design. It oversamples children living in administrative areas characterised by higher socioeconomic deprivation and larger ethnic minority population \citep{Plewis2007}. We rebalance the MCS sample to make it nationally representative by excluding from the analysis a fraction of observations from the oversampled areas, proportionally to their sampling probability.\footnote{See Table 5.5 in \citet{Plewis2007}. This choice is mainly driven by software limitations. The \texttt{lavaan} package in \textbf{\textsf{R}} \citep{Rosseel2012} is the most suitable tool for our invariance analysis, but it does not allow to use weights when outcomes are categorical, as it is the case for the socio-emotional measurements.}
We also restrict our sample to non-multiparous individuals born in England. Finally, we restrict the sample to cases where there is complete information on socioemotional skills at five years of age. The final sample contains 9,545 individuals from the British Cohort Study, and 5,572 from the Millennium Cohort Study. Summary statistics for the full and estimation samples are displayed in \autoref{tab:sumstats}. After the rebalancing step, the MCS estimation sample closely mirrors the full sample in terms of average observable characteristics, thus preserving representativeness.

@@

A common approach to infer a child's socio-emotional development is based on behavioural screening scales. As part of these tools, mothers (or teachers) indicate whether their children exhibit a series of behaviours -- the \emph{items} of the scale. In the British and Millennium Cohort Studies, two different scales were employed. In the BCS, the Rutter A Scale was used \citep{Rutter1970} while in the MCS cohort, mothers were  administered the Strengths and Difficulties Questionnaire (SDQ, \citealp{Goodman1994,Goodman1997}). The SDQ was created as an update to the Rutter scale. It encompasses more recent advances in child psychopathology, and emphasises positive traits alongside undesirable ones \citep{Stone2010}. The Rutter and SDQ scales are reproduced in \autoref{tab:scales}; they have 23 and 25 items each, respectively. In the child psychiatry and psychology literatures, the Rutter and SDQ behavioural screening scales are regarded as measures of behavioural problems and mental health. However, in our analysis we follow the economics literature, and - after having recoded them accordingly - we interpret them as measures of positive child development \citep{Goodman2011}.

While the Rutter and SDQ scales are similar in their components, there is no a priori reason to expect them to be directly comparable. First, the overlap of behaviours described in the two scales is only partial. Second, the wording of each item is slightly different, both in the description and in the options that can be selected as answers. Third, the different ordering of the items within each scale might lead to order effects. Fourth, and no less importantly, the interpretation of each behaviour by respondents living 30 years apart (1975 vs 2006) might differ due to a host of evolving societal norms.

As our goal is to compare socio-emotional skills across the two cohorts, we construct a new scale by retaining the items that are worded in a similar way across the two original Rutter and SDQ scales, and making some slight coding adjustments to maximise comparability. In what follows, we will consider the included items to be the same \emph{measure} in the two cohorts. The wording of the items we will be using in the analysis is presented in \autoref{tab:scalecomp}: we retain 13 items for the BCS (two of them are grouped) and 11 for the MCS with high degree of comparability. We exclude from the analysis items that were completely different between the two questionnaires, although we could have included them in the factor analysis and treated them as missing in the cohort were they were not administered. While this could have improved efficiency, we decided to rely on a more coherent set of measures to maximise comparability between the two cohorts. More details on the derivation of the scale are available in Appendix \autoref{asec:scales}.

@@

The original Rutter scale, used for the BCS cohort, distinguishes behaviours into two subscales: \emph{anti-social} and \emph{neurotic} \citep{Rutter1970}. This two-factor conceptualisation has been validated using data from multiple contexts, and the latent dimensions have been broadly identified as externalising and internalising behaviour problems.\footnote{See for example \citet{Fowler1979,Venables1983,Tremblay1987,Berglund1999,Klein2009}. However, in some cases a three-factor structure was found to better fit the data, with the externalising factor separating into two factors seemingly capturing aggressive and hyperactive behaviours \citep{Behar1974,McGee1985}.} The Strength and Difficulties Questionnaire, used for the MCS cohort, was conceived to have five subscales of five items each. The five subscales are: \emph{hyperactivity}, \emph{emotional symptoms}, \emph{conduct problems}, \emph{peer problems}, and \emph{prosocial}. This five-factor structure has been validated in many contexts \citep{Stone2010}; lower-dimensional structures have been also suggested \citep{Dickey2004}. Recent research has shown that there are some benefits to using broader subscales that correspond to the externalising and internalising factors in Rutter, especially in low-risk or general population samples \citep{Goodman2010}.

We use exploratory factor analysis (EFA) to assess the factors structure of our 11-item scale combining Rutter and SDQ.\footnote{Factor-analytic methods have long been used in psychology, and in recent years they have become increasingly popular in economics, especially to meaningfully aggregate high-dimensional items measuring different aspects of common underlying dimensions of human development. The EFA is performed decomposing the polychoric correlation matrix of the items and using weighted least squares, and the solution is rescaled using oblique factor rotation (\emph{oblimin}). We use the \textbf{\textsf{R}} package \texttt{psych}, version 1.8.4 \citep{Revelle2018}.} We start by investigating the number of latent constructs that are captured by the scale, using different methods developed in the psychometric literature, and recently adopted by the economics literature. The results are displayed in \autoref{tab:numfac}. As pointed out in \citet{Conti2014}, there is relatively little agreement among procedures; this is the case especially for the Rutter items in the BCS data, where different methods suggest to retain between 1 and 3 factors, while most methods suggest to retain 2 factors in the MCS. In our analysis, we adopt two factors and a dedicated measurement system, where each measure reflects only one factor. This choice is justified both by the child psychology literature cited above, and as compromise to work with the same number of factors in the two cohorts.

The two-factor EFA delivers a neat and sensible separation between items, as shown in \autoref{tab:load}: reassuringly, similarly-worded items load on the same factor across the two cohorts, and also the magnitude of the respective loadings (measuring the strength of the association between the item and the factor) is very similar. Following previous research, we name the first dimension \emph{Externalising skills} (EXT, indicating low scores on the items restless, squirmy/fidgety, fights/bullies, distracted, tantrums, and disobedient) and the second dimension \emph{Internalising skills} (INT, indicating low scores on the items worried, fearful, solitary, unhappy, and aches).\footnote{Internalising and externalising dimensions emerge from the exploratory step on our novel 11-item scale. Appendix \autoref{asec:fullefa} performs the same exploratory steps on the full set of Rutter items in BCS and SDQ items in MCS. It confirms that the items we consider in our subscale have a broadly consistent covariance structure even when considered in the context of their original scales. Appendix \autoref{asec:no511} considers the robustness of our results to the exclusion of the items of the scale that perform most poorly.}

@@

We make the common assumption of a dedicated (or congeneric) factor structure, where each measure is assumed to load on only one latent dimension \citep{Heckman2013,Conti2010a,Attanasio2018a}. We mirror the structure found in the exploratory factor analysis (see \autoref{tab:load}), and assume that items 1-6 load exclusively on the externalising factor and items 7-11 on the internalising factor.\footnote{The dedicated factor structure corresponds to a sparse loading matrix, i.e.: $$\bm{\Lambda}_{c} \coloneqq \begin{bmatrix} \lambda_{1c}, \dots, \lambda_{6c} & \bm{0} \\ \bm{0} & \lambda_{7c}, \dots, \lambda_{11c} \end{bmatrix}.$$}

@@

Measurement invariance analysis necessarily starts starts from a minimally restrictive model, denominated \emph{configural} model. This is a `minimum' identifiable model, in that it places the least possible restrictions on how parameters are allowed to vary across cohorts. It serves as a basis for our measurement invariance analysis.

@@

\item \underline{Anchored parameterisation} {[MT]} \citep{Millsap2004}
  \bi
  \item For all groups, normalise a reference loading to 1 for each factor
  \item Set invariant across groups one threshold per item (e.g. $\tau_{0,Ai} = \tau_{0,Bi}$), and an additional threshold in the reference items above
  \item In the first group: $\bm{\kappa_A}=\bm{0}$, $\text{diag}(\bm{\Sigma}_A)=\bm{I}$
  \item Set all intercepts $\bm{\nu}$ to zero
  \ei
\ei
The first two parameterisations (WE{\textDelta} and WE{\textTheta}) normalise the mean and variance of factors to the same constants in both groups, and they leave all loadings and intercepts to be freely estimated; they only differ in whether the additional required normalisation is imposed on the variances of the error terms ($\bm{\Psi}$) or on the diagonal of the covariance matrix of the measures ($\bm{\Sigma}$). The MT parameterisation instead proceeds by identifying parameters in one group first, and then imposing cross-group equality constraints to identify parameters in other groups \citep{Wu2016a}. Still, all of these parameterisations are statistically equivalent. The measurement invariance analysis in this paper is based on the Theta parameterisation (WE{\textTheta}), but results are independent on this choice. The restrictions in \eqref{eq:propens}, \eqref{eq:thresh}, and \eqref{eq:thetapar} define the so-called \emph{configural} model.

@@

Comparison of $\chi^2$ values across models is a common likelihood-based strategy. However, tests based on $\Delta \chi^2$ are known to display high Type I error rates with large sample size and more complex models such as our own \citep{Sass2014}. In fact, for all invariance levels in our applications a chi-squared difference would point to a lack of measurement invariance. The use of approximate fit indices (AFIs) is therefore recommended alongside $\chi^2$. While these indices successfully adjust for model complexity \citep{Cheung2002}, they do not have a known sampling distribution. This makes it necessary to rely on simulation studies, which derive rules of thumb indicating what level of \textDelta AFI is compatible with invariances invariance.

Again, just like in the broader context of measurement invariance, most evidence regarding the performance of AFIs pertains to scenarios with continuous measures. The root mean squared error of approximation (RMSE) and the Tucker-Lewis index (TLI) are traditionally the most used AFIs in empirical practice. Simulation evidence by \citet{Cheung2002} shows that these indices can show correlation between overall and relative fit, and suggest relying on additional indices, such as the comparative fit index (CFI, \citealp{Bentler1990}), McDonald non-centrality index (MFI, \citealp{McDonald1989}), and Gamma-hat index \citep{Steiger1989}. Subsequent simulation studies -- e.g. \citet{Chen2007} and \citet{Meade2008} -- have updated these thresholds for the continuous case. In particular, \citet{Chen2007} suggests the following thresholds for rejecting measurement invariance: \textDelta RMSE $> .015$, \textDelta CFI $< - .010$, \textDelta SRMR $> .010$.

It is not advisable to directly extrapolate rules of thumb derived from simulations with continuous measures to the categorical case \citet{Lubke2004}. Recent studies have advanced simulation evidence around the performance of AFIs in measurement invariance analysis with categorical measures. \citet{Sass2014} finds that the cutoffs from \citet{Chen2007} might not generalise well to problems estimated by WLSMV, but this is mostly confined to smaller sample sizes and detection of small degrees of non-invariance. More recently, \citet{Rutkowski2017} find that a \textDelta RMSE threshold of $.010$ is appropriate for testing equality of slopes and thresholds when the sample size is large, like in our case.

In any case, we present a range of fit indices to provide a more holistic assessment of measurement invariance. The fit of each nested model is compared in Panel A of \autoref{tab:fit}. The overall fit of the configural model is satisfactory according to all indices, with CFI around $.95$ and RMSE just above $.05$. As expected, given our large sample size, $\chi^2$-based tests reject measurement invariance at all levels. The model with restricted thresholds and loadings exhibits a comparable fit to the configural model, according to all the AFIs. In particular, the \textDelta AFIs fall within the ranges suggested in \citet{Chen2007} and \citet{Rutkowski2017}. Invariance of loadings and thresholds across cohorts implies that items in our scale are equally salient in their informational content, and that the latent propensities have equal mapping into the observed items.

@@

This is an important finding, which has to our knowledge never been acknowledged in this literature. How can this lack of comparability be explained?  A possible interpretation is connected with secular evolution of social and cultural norms about child behaviours. For example, commonly held views of what constitutes a restless, distracted, or unhappy child might have changed between 1975 and 2006.\footnote{Calibrating the Rutter and SDQ using a contemporary sample of children cannot rule out this issue. For example, \citet{Collishaw2004} administered both Rutter and SDQ items to parents of a small sample of adolescents in London. They use the mapping between the two questionnaires to impute Rutter scores for mothers who answered the SDQ. This can correct for contemporaneous reporting differences between questionnaires, but cannot tackle reporting differences between samples collected at different times in history.}

To summarise, our measurement invariance analysis shows partial comparability of socioemotional skills across cohorts. In particular, the variance of skills can be compared across cohorts, but mean cohort differences do not necessarily reflect differences in skills. We can use scores from our scale to compare children within the same cohort, but not across cohorts. However, we can also compare within-cohort differences between groups of children, across cohorts. As an example, consider two groups of children A and B in the BCS cohort, and two groups of children C and D in the MCS. We cannot compare the mean level of skills between groups A and C, but we can compare the mean difference between groups A and B with the mean difference between groups C and D. This is the approach we take for the rest of the paper. Refraining from direct cross-cohort comparisons, we interpreting significance and magnitude of within-cohort differences across the cohorts.

>>>>

Another complication arises from the fact that, differently from the British Cohort Study, the Millennium Cohort Study has a stratified design. It oversamples children living in administrative areas characterised by higher socioeconomic deprivation and larger ethnic minority population \citep{Plewis2007}. We rebalance the MCS sample to make it nationally representative by excluding from the analysis a fraction of observations from the oversampled areas, proportionally to their sampling probability.\footnote{See Table 5.5 in \citet{Plewis2007}. This choice is mainly driven by software limitations. The \texttt{lavaan} package in \textbf{\textsf{R}} \citep{Rosseel2012} is the most suitable tool for our invariance analysis, but it does not allow to use weights when outcomes are categorical, as it is the case for the socio-emotional measurements.} Finally, we also restrict our sample to individuals born in England and to cases where there is complete information on socioemotional skills at five years of age. The final sample contains 9,545 individuals from the British Cohort Study, and 5,572 from the Millennium Cohort Study. Summary statistics for the full and estimation samples are displayed in \autoref{tab:sumstats}. After the rebalancing step, the MCS estimation sample closely mirrors the full sample in terms of average observable characteristics, thus preserving representativeness.

@@

A common approach to infer a child's socio-emotional development is based on behavioural screening scales. As part of these tools, mothers (or teachers) indicate whether their children exhibit a series of behaviours -- the \emph{items} of the scale. In the British and Millennium Cohort Studies, two different scales were employed. In the BCS, the Rutter A Scale was used \citep{Rutter1970} while in the MCS cohort, mothers were  administered the Strengths and Difficulties Questionnaire (SDQ, \citealp{Goodman1994,Goodman1997}). The SDQ was created as an update to the Rutter scale. It encompasses more recent advances in child psychopathology, and emphasises positive traits alongside undesirable ones \citep{Stone2010}. \citet{Goodman1997} administered both scales to a sample of children, and showed that the scores are highly correlated, and the two measures do not differ in their discriminatory ability. The Rutter and SDQ scales are reproduced in \autoref{tab:scales}; they have 23 and 25 items each, respectively. In the child psychiatry and psychology literatures, the Rutter and SDQ behavioural screening scales are regarded as measures of behavioural problems and mental health. However, in our analysis we follow the economics literature, and - after having recoded them accordingly - we interpret them as measures of positive child development \citep{Goodman2011}.

While the Rutter and SDQ scales are similar in their components (since the latter was developed from the former, see \citet{Goodman1994}), there is no a priori reason to expect them to be directly comparable. First, the overlap of behaviours described in the two scales is only partial, given that - by design - the SDQ includes also strengths, in addition to weaknesses. Second, the wording of each item is slightly different, both in the description and in the options that can be selected as answers. Third, the different ordering of the items within each scale might lead to order effects. Fourth, and no less importantly, the interpretation of each behaviour by respondents living 30 years apart (1975 vs 2006) might differ due to a host of evolving societal norms. Nonetheless, the level of comparibility of the two scales is higher than that of other scales used in comparating work in the literature reviewed in section \ref{sec:lit}.

As our goal is to compare socio-emotional skills across the two cohorts, we construct a new scale by retaining the items that are worded in a similar way across the two original Rutter and SDQ scales, and making some slight coding adjustments to maximise comparability. In what follows, we will consider the included items to be the same \emph{measure} in the two cohorts. The wording of the items we will be using in the analysis is presented in \autoref{tab:scalecomp}: we retain 13 items for the BCS (two of them are grouped) and 11 for the MCS with high degree of comparability. We exclude from the analysis items that were completely different between the two questionnaires to maximise comparability between the two cohorts, as it is standard good practice in the psychometric literature (see for example \citet{Kern2014}).\footnote{Of course, we could have included them in the factor analysis and treated them as missing in the cohort were they were not administered.} More details on the derivation of the scale are available in Appendix \autoref{asec:scales}.

@@

The original Rutter scale, used for the BCS cohort, distinguishes behaviours into two subscales: \emph{anti-social} and \emph{neurotic} \citep{Rutter1970}. This two-factor conceptualisation has been validated using data from multiple contexts, and the latent dimensions have been broadly identified as externalising and internalising behaviour problems.\footnote{See for example \citet{Fowler1979,Venables1983,Tremblay1987,Berglund1999,Klein2009}. However, in some cases a three-factor structure was found to better fit the data, with the externalising factor separating into two factors seemingly capturing aggressive and hyperactive behaviours \citep{Behar1974,McGee1985}.} The Strength and Difficulties Questionnaire, used for the MCS cohort, was conceived to have five subscales of five items each. The five subscales are: \emph{hyperactivity}, \emph{emotional symptoms}, \emph{conduct problems}, \emph{peer problems}, and \emph{prosocial}. This five-factor structure has been validated in many contexts \citep{Stone2010}; lower-dimensional structures have been also suggested \citep{Dickey2004}. Recent research has shown that there are some benefits to using broader subscales that correspond to the externalising and internalising factors in Rutter, especially in low-risk or general population samples \citep{Goodman2010}. Indeed, the internalising and externalising dimensions were introduced in psychology by \citet{Achenbach1966}, who show that they are the two main factors underlying a wide range of psychological measures; as noted in \citet{Achenbach2016}, more than 75,000 articles have been published on internalizing and externalizing problems.

We use exploratory factor analysis (EFA) to assess the factor structure of our new scale, composed of 11 items of the Rutter scale in the BCS and the corresponding items of the SDQ in the MCS.\footnote{Factor-analytic methods have long been used in psychology, and in recent years they have become increasingly popular in economics, especially to meaningfully aggregate high-dimensional items measuring different aspects of common underlying dimensions of human development. The EFA is performed decomposing the polychoric correlation matrix of the items and using weighted least squares, and the solution is rescaled using oblique factor rotation (\emph{oblimin}). We use the \textbf{\textsf{R}} package \texttt{psych}, version 1.8.4 \citep{Revelle2018}.} We start by investigating the number of latent constructs that are captured by the scale, using different methods developed in the psychometric literature, and recently adopted by the economics literature. The results are displayed in \autoref{tab:numfac}. As pointed out in \citet{Conti2014}, there is relatively little agreement among procedures; this is the case especially for the Rutter items in the BCS data, where different methods suggest to retain between 1 and 3 factors, while most methods suggest to retain 2 factors for the SDQ items in the MCS. In our analysis, we adopt two factors, which is the average number of factors across all methods for each cohort and gender. This choice is also consistent with the child psychology literature cited above, and allows us to work with the same number of factors in the two cohorts.

The two-factor EFA delivers a neat and sensible separation between items, as shown in \autoref{tab:load}: reassuringly, similarly-worded items load on the same factor across the two cohorts, and also the magnitude of the respective loadings (measuring the strength of the association between the item and the factor) is very similar. Following previous research, we name the first dimension \emph{Externalising skills} (EXT, indicating low scores on the items restless, squirmy/fidgety, fights/bullies, distracted, tantrums, and disobedient) and the second dimension \emph{Internalising skills} (INT, indicating low scores on the items worried, fearful, solitary, unhappy, and aches).\footnote{Internalising and externalising dimensions emerge from the exploratory step on our novel 11-item scale. Appendix \autoref{asec:fullefa} performs the same exploratory steps on the full set of Rutter items in BCS and SDQ items in MCS. It confirms that the items we select for our subscale have a broadly consistent covariance structure even when factor-analysed with the others in their original scales. Appendix \autoref{asec:no511} considers the robustness of our results to the exclusion of the items of the scale that perform most poorly.}

@@

We make the common assumption of a dedicated (or congeneric) factor structure, where each measure is assumed to load on only one latent dimension \citep{Heckman2013,Conti2010a,Attanasio2018a}. We mirror the structure found in the exploratory factor analysis (see \autoref{tab:load}), and assume that items 1-6 load exclusively on the externalising factor and items 7-11 on the internalising factor.\footnote{The dedicated factor structure corresponds to a sparse loading matrix, i.e.: $$\bm{\Lambda}_{c} \coloneqq \begin{bmatrix} \lambda_{1c}, \dots, \lambda_{6c} & \bm{0} \\ \bm{0} & \lambda_{7c}, \dots, \lambda_{11c} \end{bmatrix}.$$} This choice of using a dedicated measurement system is consistent with the standard rule of thumb for selecting items based on the magnitude of their loadings, which is indicated as 0.32 by \citet{Tabachnick2001}.

@@

Measurement invariance analysis necessarily starts from a minimally restrictive model, denominated \emph{configural} model. This is a `minimum' identifiable model, in that it places the least possible restrictions on how parameters are allowed to vary across cohorts. It serves as a basis for our measurement invariance analysis.

@@

\item \underline{Anchored parameterisation} {[MT]} \citep{Millsap2004}
  \bi
  \item For all groups, normalise a reference loading to 1 for each factor
  \item Set invariant across groups one threshold per item (e.g. $\tau_{0,Ai} = \tau_{0,Bi}$ for two groups $A$ and $B$), and an additional threshold in the reference items above
  \item In the first group: $\bm{\kappa_A}=\bm{0}$, $\text{diag}(\bm{\Sigma}_A)=\bm{I}$
  \item Set all intercepts $\bm{\nu}$ to zero
  \ei
\ei
The first two parameterisations (WE{\textDelta} and WE{\textTheta}) normalise the mean and variance of factors to the same constants in both groups, and they leave all loadings and thresholds to be freely estimated; they only differ in whether the additional required normalisation is imposed on the variances of the error terms ($\bm{\Psi}$) or on the diagonal of the covariance matrix of the measures ($\bm{\Sigma}$). The MT parameterisation instead proceeds by identifying parameters in one group first, and then imposing cross-group equality constraints to identify parameters in other groups \citep{Wu2016a}. Still, all of these parameterisations are statistically equivalent. The measurement invariance analysis in this paper is based on the Theta parameterisation (WE{\textTheta}), but results are independent on this choice. The restrictions in \eqref{eq:propens}, \eqref{eq:thresh}, and \eqref{eq:thetapar} define the so-called \emph{configural} model.

@@

Comparison of $\chi^2$ values across models is a common likelihood-based strategy. However, tests based on $\Delta \chi^2$ are known to display high Type I error rates with large sample size and more complex models such as our own \citep{Sass2014}. In fact, for all invariance levels in our applications a chi-squared difference would point to a lack of measurement invariance. The use of approximate fit indices (AFIs) is therefore recommended alongside $\chi^2$. While these indices successfully adjust for model complexity \citep{Cheung2002}, they do not have a known sampling distribution. This makes it necessary to rely on simulation studies, which derive rules of thumb indicating what level of \textDelta AFI is compatible with invariance.

Again, just like in the broader context of measurement invariance, most evidence regarding the performance of AFIs pertains to scenarios with continuous measures. The root mean squared error of approximation (RMSE) and the Tucker-Lewis index (TLI) are traditionally the most used AFIs in empirical practice. Simulation evidence by \citet{Cheung2002} shows that these indices can show correlation between overall and relative fit, and suggest relying on additional indices, such as the comparative fit index (CFI, \citealp{Bentler1990}), McDonald non-centrality index (MFI, \citealp{McDonald1989}), and Gamma-hat index \citep{Steiger1989}. Subsequent simulation studies -- e.g. \citet{Chen2007} and \citet{Meade2008} -- have updated these thresholds for the continuous case. In particular, \citet{Chen2007} shows in two Monte Carlo studies that the standardised root mean square residual (RMSR) is more sensitive to lack of invariance in factor loadings than in intercepts or residual variances, while the CFI and RMSEA are equally sensitive to all three types of lack of invariance; he suggests the following thresholds for rejecting measurement invariance: \textDelta RMSE $> .015$, \textDelta CFI $< - .010$, \textDelta SRMR $> .010$.

However, it is not advisable to directly extrapolate rules of thumb derived from simulations with continuous measures to the categorical case \citep{Lubke2004}. Recent studies have advanced simulation evidence around the performance of AFIs in measurement invariance analysis with categorical measures. \citet{Sass2014} finds that the cutoffs from \citet{Chen2007} might not generalise well to problems estimated by WLSMV, but this is mostly confined to smaller sample sizes and detection of small degrees of non-invariance. More recently, \citet{Rutkowski2017} find that a \textDelta RMSE threshold of $.010$ is appropriate for testing equality of slopes and thresholds when the sample size is large, like in our case.

In any case, we present a range of fit indices to provide a more holistic assessment of measurement invariance. The fit of each nested model is compared in Panel A of \autoref{tab:fit}. The overall fit of the configural model is satisfactory according to all indices, with CFI around $.95$ and RMSE just above $.05$. As expected, given our large sample size, $\chi^2$-based tests reject measurement invariance at all levels. The model with restricted thresholds and loadings exhibits a comparable fit to the configural model, according to all the AFIs. In particular, the \textDelta AFIs fall within the ranges suggested in \citet{Chen2007}, \citet{Rutkowski2017} and \citet{Svetina2017}; see also \citet{Svetina2019} for a review of updated guidelines for measurement invariance. Invariance of loadings and thresholds across cohorts implies that items in our scale are equally salient in their informational content, and that the latent propensities have equal mapping into the observed items.

@@

This is an important finding, which has to our knowledge never been acknowledged in the economic literature. How can this lack of comparability be explained?  A possible interpretation is connected with secular evolution of social and cultural norms about child behaviours. For example, commonly held views of what constitutes a restless, distracted, or unhappy child might have changed between 1975 and 2006.\footnote{Calibrating the Rutter and SDQ using a contemporary sample of children cannot rule out this issue. For example, \citet{Collishaw2004} administered both Rutter and SDQ items to parents of a small sample of adolescents in London. They use the mapping between the two questionnaires to impute Rutter scores for mothers who answered the SDQ. This can correct for contemporaneous reporting differences between questionnaires, but cannot tackle reporting differences between samples collected at different times in history.}

To summarise, our measurement invariance analysis shows partial comparability of socioemotional skills across cohorts. In particular, the variance of skills can be compared across cohorts, but mean cohort differences do not necessarily reflect differences in skills. We can use scores from our scale to compare children within the same cohort-gender group, but not across cohorts. However, we can also compare within-cohort differences between groups of children, across cohorts. As an example, consider two groups of children A and B in the BCS cohort, and two groups of children C and D in the MCS. We cannot compare the mean level of skills between groups A and C, but we can compare the mean difference between groups A and B with the mean difference between groups C and D. This is the approach we take for the rest of the paper. Refraining from direct cross-cohort comparisons, we interpreting significance and magnitude of within-cohort differences across the cohorts.
